---
title: "Reproducible Research: Peer Assessment 1"
author: "Tom Bruning"
output: 
html_document:
keep_md: true
---

### Initial setup

```{r required packages, warning=FALSE, message=FALSE}
require(dplyr)
require(tidyr)
require(ggplot2)
require(data.table)
require(xtable)
require(lubridate)
require(ggvis)
options(scipen = 1)
```

### Loading and preprocessing the data
Since I am using a static file, downloaded and unzipped once I am not including downloading the file step.  
I use the data table package since running in tables is faster than running as data frames.
```{r read table}
DT <- read.table("activity.csv", sep = ",", header = TRUE)
```
Create time interval using hour and minute
```{r create time}
DT1 <- DT %>% 
    mutate(hour = floor(interval / 100)) %>% 
    mutate(min = ifelse(hour < 1, interval, interval - 100 * hour)) %>%
    mutate(hour1 = ifelse(hour < 10, paste0("0", hour, sep = ""), hour)) %>% 
    mutate(min1 = ifelse(min < 10, paste0("0", min, sep = ""), min)) %>% 
    mutate(time = paste(hour1, ":", min1, sep = "")) %>% 
    select(steps, date, time)
```
Delete missing values
```{r delete missing values}
DT_comp <- na.omit(DT1)
DT_comp$steps <- as.numeric((DT_comp$steps)) 
```
### What is mean total number of steps taken per day?
```{r histogram, fig.height=4, fig.width=5}
dt_summ <- 
    DT_comp %>% 
    group_by(date) %>% 
    summarise(tot = (sum(steps)))
dt_summ %>% 
    ggvis(~tot, fill := "white" ) %>% 
    layer_histograms( width = 5000, center = 7500) %>% 
    add_axis("x", title = "Steps", values = seq(0, 25000, by = 5000)) %>% 
    add_axis("y", title = "Frequency")


mean <- prettyNum(mean <- round(mean(dt_summ$tot),0), big.mark = ",")
median <- prettyNum(median <- round(median(dt_summ$tot),0), big.mark = ",")
total <- prettyNum(total <- round(sum(dt_summ$tot),0), big.mark = ",")
```

#### The mean number of steps per day is `r mean`. The median number of steps per day is `r median`, and the total number of steps is `r total`.

### What is the average daily activity pattern?
``` {r, pattern,fig.height=3, fig.width=4}
DT_comp %>% 
    group_by(time)
dt_summ_int <- 
    DT_comp %>% 
    group_by(time) %>% 
    summarise(mean = round((mean(steps)), 0))
dt_summ_int %>% 
    ggvis(~time, ~mean) %>% 
    layer_lines(stroke := "red")
```

### Imputing missing values

```{r missing_values, fig.height=3, fig.width=4}
# Calculate number of missing values.
y <- which(is.na(DT1)==TRUE)
y_len <-  prettyNum(length(y), big.mark = ",")
# Create summary of the mean of the data by time intervals
dt_summ <- 
    DT_comp %>% 
    group_by(time) %>% 
    summarise(mean = mean(steps))
# Create copy of original data and populate NA's with meaan of the interval over the data set.
DT2 <- DT1
DT2$steps[y] <- round(dt_summ$mean[match(dt_summ$time, DT2$time)],0)
dt_summ_new <- 
    DT2 %>% 
    group_by(date) %>% 
    summarise(tot = sum(steps))
# Create new histogram
dt_summ_new %>% 
    ggvis(~tot, fill := "white" ) %>% 
    layer_histograms( width = 5000, center = 7500) %>% 
    add_axis("x", title = "Steps", values = seq(0, 25000, by = 5000)) %>% 
    add_axis("y", title = "Frequency")
```

`

```{r, results="asis"}
dt_summ <- 
    DT_comp %>% 
    group_by(date) %>% 
    summarise(tot = (sum(steps)))
tot <- sum(dt_summ$tot)
mean <- mean(dt_summ$tot)
median <- median(dt_summ$tot)
old_data <- c(mean, median, tot)
new_tot <- sum(dt_summ_new$tot)
new_mean <- round(mean(dt_summ_new$tot), 0)
new_median <- median(dt_summ_new$tot)
rnames <- tbl_dt(c("mean", "median", "total"))
old_data <- c(mean, median, tot)
new_data <- c(new_mean, new_median, new_tot)
new_data <- tbl_dt(new_data)
old_data <- tbl_dt(old_data)
rnames <- tbl_dt(c("mean", "median", "total"))
new_tbl <- bind_cols(rnames, old_data, new_data)
names(new_tbl) <- c("Value", "Omitted", "Imputed")
# new_tbl <- as.data.frame(new_tbl)
ptable <- xtable(new_tbl, format = "html", digits=0, caption = c("Comparision of omitted vs. imputted values"))

```
The number of NA's in the original data set is `r y_len`.
new mean `r new_mean`, new median `r new_median`, new total `r new_tot
``` {r, results="asis"}
print(ptable, type = "html", include.rownames = FALSE)
```

##### To replace the omitted data I used the mean data collected for each interval.  When placing this data in the data table it doesn't affect the mean at all, it only affects the median slightly, but it has a huge effect on the total number of steps taken over the period.  Adding a small data value to the missing values (`r y_len` NA's) increases the total number of steps taken.

### Are there differences in activity patterns between weekdays and weekends?

