---
title: "Reproducible Research: Peer Assessment 1"
author: "Tom Bruning"
output: 
  html_document:
    keep_md: true
---

### Initial setup

```{r required packages, warning=FALSE, message=FALSE}
require(dplyr)
require(tidyr)
require(ggplot2)
require(data.table)
require(xtable)
options(scipen = 1)
```

### Loading and preprocessing the data
Since I am using a static file, downloaded and unzipped once I am not including downloading the file step.  
I use the data table package since running in tables is faster than running as data frames.
```{r read table}
DT <- read.table("activity.csv", sep = ",", header = TRUE)
```
Create time interval using hour and minute
```{r create time}
DT1 <- DT %>% 
    mutate(hour = floor(interval / 100)) %>% 
    mutate(min = ifelse(hour < 1, interval, interval - 100 * hour)) %>%
    mutate(hour1 = ifelse(hour < 10, paste0("0", hour, sep = ""), hour)) %>% 
    mutate(min1 = ifelse(min < 10, paste0("0", min, sep = ""), min)) %>% 
    mutate(time = paste(hour1, ":", min1, sep = "")) %>% 
    select(steps, date, time)
glimpse(DT1)
```
Delete missing values
```{r delete missing values}
DT_comp <- na.omit(DT1)
DT_comp$steps <- as.numeric((DT_comp$steps)) 
glimpse(DT_comp)
```
### What is mean total number of steps taken per day?
```{r histogram}
dt_summ <- 
    DT_comp %>% 
    group_by(date) %>% 
    summarise(tot = (sum(steps)))
hist(dt_summ$tot,
     xlab ="Steps",
     main = "Step per Day"
     )
mean <- prettyNum(mean <- round(mean(dt_summ$tot),0), big.mark = ",")
median <- prettyNum(median <- round(median(dt_summ$tot),0), big.mark = ",")
```

#### The mean number of steps per day is `r mean`. The median number of steps per day is `r median`.

### What is the average daily activity pattern?
```{r daily}


```





### Imputing missing values



### Are there differences in activity patterns between weekdays and weekends?

